---
title: "A list of ggplot2 tips for quality plots"
author: RF
---

```{r}
library(ggplot2)
library(dplyr)
library(mizer)
```



There shall be no order in this markdown, just an endless succession of bits of code to plot nice things

Things to know:

- Format your dataframe correctly before starting, this is 90% of the work (help: reshape2::melt())
- Check units within the dataframe. Numerics for continous var and factors for discretes ones

- if you cannot do something in ggplot2, you probably shouldn't do it for a paper plot (looking at you double y axis)

- ggplot2 is like an onion (or Shrek), just layers over layers.

What datasets are we going to play with?

```{r}
data()
```


## Aesthetics 

Var within `aes()` are taken from the df and added to the legend. Var outside are constants and do not appear in legend

## Legends

To overwrite colors legends

```{r}
scale_color_manual(values = c("red","black")) 
```

Can give a vector with a name, this will overwrite legend name too. "color" because we are coloring a line, but if it was a box plot it would be "fill" for example. "manual" because I decide which colors but it could be an automatic function such as "gradient".

## Themes

Some theme are already customized such as `theme_bw())` or `theme_minimal()`
I like to write my own theme. Here are some examples

```{r, eval = FALSE}
legend.title=element_text(),
panel.background = element_rect(fill = "white", color = "black"), # black for the contour, white for the background
legend.position = "bottom", # self explanatory
strip.background = element_blank(), # for panels, strip is the box with the var name. I don't like the grey
legend.key = element_rect(fill = "white")) # background of the color/symbol in the legend. Still don't like grey
```



## Texts

## Facets

grid: good for looking at 2 var

wrap: good to look at many facets grouped by one var

## Guides

Arguments in `guides()` take function named `guide_something()`

e.g.

```{r}
guides(color = guide_legend(nrow=2))
```

To organise the legend to be displayed in 2 rows


## Randos

`coord_flip()`


## geom_ribbon

Ribbon take a ymin and ymax. One can use the standard deviation of a gam for geom_smooth (which is ugly) to create a personalised ribbon. Below non-working example because missing data

```{r}
# prepare the data for the ribbon (standard deviation)
g1 <- ggplot(plot_dat[[1]])+
  geom_smooth(aes(x = time, y = meanSp-sdSp, group = as.factor(sp))) +
  geom_smooth(aes(x= time, y = meanSp+sdSp, group = as.factor(sp))) 
gg1 <- ggplot_build(g1)
dfRibbon1 <- data.frame(x = gg1$data[[1]]$x, ymin = gg1$data[[1]]$y, ymax = gg1$data[[2]]$y, group = gg1$data[[1]]$group) #and extract the smooth data for the ribbon

p1 <- ggplot(plot_dat[[1]])+
  geom_line(aes(x = time, y = meanSp, color = as.factor(sp)))+
  geom_ribbon(data = dfRibbon1, aes(x = x, ymin = ymin, ymax = ymax, group = group), fill = "grey", alpha = 0.2)

```


# geom_pointrange

```{r}

edf <- readRDS("data/anovaDat.rds")

#backtransform to orginal eta scale
edf$fit <-exp(edf$fit)
edf$lower <-exp(edf$lower)
edf$upper <-exp(edf$upper)

## plot effects
edf$etaMean <-edf$fit

# formating for nice plot
edf$predation <- as.character(edf$predation)
edf$fisheries <- as.character(edf$fisheries)
edf$predation[which(edf$predation == 0)] <- "a) without predation"
edf$predation[which(edf$predation == 1)] <- "b) with predation"
edf$fisheries[which(edf$fisheries == 0)] <- "un-fished"
edf$fisheries[which(edf$fisheries == 1)] <- "fished"


fie <- readRDS(file = "data/anovaWeighted.rds") #weighted species mean (weighted by phenotype abundance) at year 6000
fie$run <- fie$simulation
fie$species <- as.factor(fie$species)
fie$predation[which(fie$predation == 0)] <- "a) without predation"
fie$predation[which(fie$predation == 1)] <- "b) with predation"
fie$fisheries[which(fie$fisheries == 0)] <- "un-fished"
fie$fisheries[which(fie$fisheries == 1)] <- "fished"


ggplot(edf,aes(x=species,y=etaMean))  +
  # geom_point(data=fie,color="dark grey",size=0.5,alpha=0.8)  +   
  facet_wrap(~predation)+
  geom_pointrange(data=edf,aes(color=fisheries,ymin=lower,ymax=upper),position=position_dodge(width=.1),alpha=0.8) +
  # scale_color_manual(values = c("red","black")) +
  # xlab("Species") + ylab("Eta") + 
  # geom_hline(yintercept=0.25, linetype="dashed", color="grey")  +
  # ylab(bquote(eta))  +
  theme(
    panel.background = element_rect(fill = "white", color = "black"),
    legend.position = "bottom",
    strip.background = element_blank(),
    legend.key = element_rect(fill = "white"))

# ggsave(p, file = "figures/Fig3.eps", units = "cm", width = 20, height = 15, device=cairo_ps)
```
# geom_errorbarh

```{r}

myDat <- readRDS("data/maxSize.rds")

# Averaging across species/sim/scenario
plot_dat <- NULL
for(iSpecies in unique(myDat$species))
{
  for(iScenario in unique(myDat$scenario))
  {
    for(iFish in unique(myDat$fisheries))
    {  
      temp <- filter(myDat, species == iSpecies & scenario == iScenario & fisheries == iFish)
      wmatMean <- mean(temp$maturationSize)
      wmatSd <- sd(temp$maturationSize)
      wmaxMean <- mean(temp$maxSizeTheoretical)
      wmaxSd <- sd(temp$maxSizeTheoretical) 
      plot_dat <- rbind(plot_dat,c(iSpecies,iScenario,iFish,wmatMean,wmatSd,wmaxMean,wmaxSd))
    }
  }
}
# formating the data
plot_dat <- as.data.frame(plot_dat)
colnames(plot_dat) <- c("species","scenario","fisheries","wmatMean","wmatSd","wmaxMean","wmaxSd")
plot_dat$wmatMean <- as.numeric(as.character(plot_dat$wmatMean))
plot_dat$wmatSd <- as.numeric(as.character(plot_dat$wmatSd))
plot_dat$wmaxMean <- as.numeric(as.character(plot_dat$wmaxMean))
plot_dat$wmaxSd <- as.numeric(as.character(plot_dat$wmaxSd))

plot_dat$scenario <- as.character(plot_dat$scenario)
plot_dat$scenario[which(plot_dat$scenario == "Sim9")] <- "predation"
plot_dat$scenario[which(plot_dat$scenario == "SimNoInter9")] <- "no predation"

plot_dat <- filter(plot_dat, fisheries != "init")
plot_dat$fisheries <- as.character(plot_dat$fisheries)
plot_dat$fisheries[which(plot_dat$fisheries == "normal")] <- "un-fished"
plot_dat$fisheries[which(plot_dat$fisheries == "fisheries")] <- "fished"

## add empirical data to the plot
empDat <- readxl::read_excel("data/OlssonGislasonDATA.xlsx")
empDat <- as.data.frame(empDat)

empDat2 <- readxl::read_excel("data/marineSizeData.xlsx")
empDat2 <- as.data.frame(empDat2)
colnames(empDat2)[1] <- "species"
empDat2$...5 <- NULL
colnames(empDat2)[2] <- "Lmax"
empDat2$wmax <- empDat2$a*empDat2$Lmax^(empDat2$b)
empDat2$wmat <- empDat2$a*empDat2$Lmat^(empDat2$b)

empDat3 <- readxl::read_excel("data/smallFish.xlsx")
empDat3 <- as.data.frame(empDat3)
# merge empirical data
empDat1 <- empDat[,c("Wmat","Winf")]
empDat2 <- empDat2[,c("wmat","wmax")]
colnames(empDat2) <- colnames(empDat1)
colnames(empDat3) <- colnames(empDat1)
empDat <- rbind(empDat1,empDat2,empDat3)

ggplot(filter(plot_dat)) +  
  geom_pointrange(aes(x = wmaxMean, y = wmatMean,ymin = wmatMean - wmatSd, ymax = wmatMean + wmatSd, color = fisheries, shape = scenario), alpha = .8, size = 1) +
  geom_errorbarh(aes(y = wmatMean,xmin = wmaxMean - wmaxSd, xmax = wmaxMean + wmaxSd, height = 1, color = fisheries)) +
  # geom_point(data = empDat, aes(x = Winf, y = Wmat), color = "grey", shape = "*", size = 10) +
  scale_x_continuous(trans = "log10", name = "Asymptotic size in g") +
  scale_y_continuous(trans = "log10", name = "Maturation size in g") +
  scale_color_manual(values = c("red","black")) +
  theme(legend.title=element_text(),
        panel.background = element_rect(fill = "white", color = "black"),
        legend.position="bottom",
        legend.key = element_rect(fill = "white"))+
  ggtitle(NULL)

# ggsave(p, file = "figures/Fig6.eps", units = "cm", width = 20, height = 20,device=cairo_ps)


```

## Violin plots

```{r}
# create a vector of PPMR bins
min_PPMR <- 10
max_PPMR <- 100000
no_bin <- 13
ppmrBin <- 10^(seq(from = log10(min_PPMR),to = log10(max_PPMR), length.out = no_bin))
ppmrBin <- ppmrBin[-c(12,13)]
dPPMR <- diff(ppmrBin)
# Correctly defined dw by using the proper ratio 
# (successive dw's have a fixed ratio). 
# dw[no_w] <- dw[no_w - 1] * (dw[no_w - 1] / dw[no_w - 2])

# Now, to put the phen in their bin
myDat <- readRDS("data/traitDat.rds")
myDat$kappa <- as.factor(myDat$kappa)
myDat$species <- as.factor(myDat$species)
kappa_vec <- c("0.05","0.1","0.2")
SpIdx <- 1:5

plot_dat <- NULL
for(ikappa in kappa_vec)
{
  temp_dat <- filter(myDat, kappa == ikappa)
  myMatrix <- matrix(0, nrow = 5, ncol = length(ppmrBin), dimnames = list("species" = 1:5, "PPMR" = ppmrBin))
  
  for(i in 1:dim(temp_dat)[1])
  {
    iSpecies <- temp_dat[i,]$species
    iPPMR <- temp_dat[i,]$beta
    iPPMR <- ppmrBin[sapply(iPPMR, function(i) which.min(abs(i - ppmrBin)))]
    myMatrix[as.character(iSpecies),as.character(iPPMR)] <- myMatrix[as.character(iSpecies),as.character(iPPMR)] + temp_dat[i,]$biomass
  }
  ##get relative biomasses of each PPMR group for each species:
  myMatrix <- myMatrix/rowSums(myMatrix)
  kappa_dat <- reshape2::melt(myMatrix)
  kappa_dat$kappa <- ikappa
  plot_dat <- rbind(plot_dat,kappa_dat)
}
# plot_dat$PPMR <- log10(plot_dat$PPMR)
plot_dat$PPMR <- round(plot_dat$PPMR)
plot_dat$PPMR <- as.factor(plot_dat$PPMR)
plot_dat$species <- as.factor(plot_dat$species)


dat_text <- data.frame(
  label = c("a)", "b)", "c)","d)","e)","f)","g)","h)","i)","j)","k)","l)","m)","n)","o)"),
  kappa   = c(0.05,.1,.2),
  species    = c(rep(1,3),rep(2,3),rep(3,3),rep(4,3),rep(5,3))
)

colfunc <- colorRampPalette(c("firebrick3","darkturquoise", "orange"))
colGrad <- colfunc(length(SpIdx))

ggplot(plot_dat) +
  geom_bar(aes(x = PPMR, y = value , fill = species), stat = "identity") +
  geom_bar(aes(x = PPMR, y = -value , fill = species), stat = "identity") +
  coord_flip() +
  scale_fill_manual(name = "species", values = colGrad)+ # color gradient
  scale_x_discrete(breaks = c(1 %o% 10^(seq(1,5))))+
  scale_y_continuous(name = "relative abundance proportion", breaks = c(0,.5))+
  facet_grid(kappa ~ species) +
  geom_text(size = 5, data = dat_text, mapping = aes(x =  Inf, y = -Inf, label = label), hjust = -0.75, vjust = 1.5) +
  theme(legend.title=element_text(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        panel.background = element_rect(fill = "white", color = "black"),
        panel.grid.minor = element_line(colour = "grey92"), 
        legend.position="bottom",
        legend.key = element_rect(fill = "white"))+
  guides(fill = guide_legend(nrow = 1))+
  ggtitle(NULL)
```


## ggrepel

```{r}
sim_guessed <- project(NS_params)
plot_dat <- as.data.frame(getRDI(sim_guessed@params)/getRDD(sim_guessed@params))
plot_dat$species <- factor(rownames(plot_dat),sim_guessed@params@species_params$species)
colnames(plot_dat)[1] <- "ratio"
plot_dat$w_inf <- sim_guessed@params@species_params$w_inf

ggplot(plot_dat)+
   geom_point(aes(x = w_inf, y = ratio, color = species), size = 6, alpha = .8) +
   ggrepel::geom_text_repel(aes(x = w_inf, y = ratio, label = species), hjust = 0, nudge_x = 0.05)+
   scale_y_continuous(name = "Density-independent / density-dependent reproduction rate", trans = "log10") +
   scale_x_continuous(name = "Asymptotic size (g)", trans = "log10") +
   scale_colour_manual(values = sim_guessed@params@linecolour) +
   theme(panel.background = element_blank(), 
         panel.border = element_rect(colour = "gray", fill=NA, size=.5),
         text = element_text(size=10),
         panel.grid.minor = element_line(color = "gray"),
         legend.position = "bottom",legend.key = element_rect(fill = "white"))


```



## maps

```{r}




```library(ozmaps)
library(tidyverse)
library(sf)
library(ggspatial)

oz_states <- ozmaps::ozmap_states

capital <- data.frame("name" = c("Hobart","Launceston"), "lat"= c(-42.8821, -41.4391), "lon"= c(147.3272, 147.1358))

### polygons

test <- data.frame(x = c(145,145.5,144.5,144), y = c(-40,-42,-41.5,-40.5))

ggplot(filter(oz_states, NAME == "Tasmania" )) +
  geom_polygon(data = test, aes(x = x, y=y), alpha = .25)+
  geom_sf(alpha = 1) +
  geom_point(data = capital, aes(x = lon, y = lat, color = name), shape = 18, size = 5,) +
  annotation_scale(location = "bl", width_hint = 0.25) +
  ggrepel::geom_text_repel(data = capital, aes(x = lon, y = lat, label = name)) +
  scale_color_manual(values = c("red","black")) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.25, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_fancy_orienteering) +
  # geom_raster(data = r.pts@data[1:20,], aes(x = long, y = lat, fill = layer))+
  geom_polygon(data = test, aes(x = x, y=y), alpha = .25)+
  theme(legend.position = "none")





